Sequential rank agreement methods for comparison of ranked lists

### 2018-02-25 latex/biosts3.tex
Updated readme with example code (simpel and more advanced)

In this article we introduce sequential rank agreement for measuring
agreement among  ranked lists.  The general idea is to
define agreement based on the sequence of ranks from a subset of the first $d$
items in each list. As agreement metric we adapt the limits of
agreement known from agreement between quantitative variables
\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
could essentially be used. Our proposed approach allows us to
compare multiple lists simultaneously, it provides a dynamic
measure of agreement as a function of the depth in the lists, it places
higher weight on items at the top of the list, it accommodates partially
observed lists of varying lengths, and has a natural interpretation
that directly relates to the ranks.  Graphical illustration of
sequential rank agreement allows us to infer any
changepoints, i.e., a list depth where a substantial change in the agreement of
the lists occur but we also provide asymptotical and randomization-based graphical
tools to compare the observed rank agreement to the expected agreement
found in non-informative data.  In this sense our approach is a
combination and generalization of some of the ideas of
\citet{Carterette2009} and \citet{Boulesteix2009}. The former compares
two rankings based on the distance between them as measured by a
multivariate Gaussian distribution and the latter presents an overview
of approaches for aggregation of ranked lists including bootstrap and
leave-one-out jackknife approaches. We show the asymptotic
distribution of the endpoint of agreement and discuss how to infer the
distribution in small-sample situations using computational
methods. This enables us to make inferences about the endpoint of
agreement even for situations where there is no actual changepoint
present but just a gradual decline in agreement.

@@

Other recent approaches consider the intersection of lists as the basis
for a similarity measure. However, simple intersection also places
equal weights on all depths of the list and therefore \citet{Fagin2003} and
\citet{Webber2010} proposed weighted intersections which put more
emphasis on the top of the lists. 
% One proposal is denoted the average overlap (AO) and used for
% comparison later on. 
Specifically, \citet{Webber2010} define their rank-biased overlap
(RBO) by weighting with a converging series to ensure that the top is
weighted higher than the potentially non-informative bottom of the
lists.  It is possible to use the existing methods to calculate
agreement of lists until a given depth, i.e., limited to the $d$ items
of each list. However, the interpretation may not be straightforward,
especially in the case of more than two lists, and they may not
accommodate partial rankings.

>>>>

In this article we introduce sequential rank agreement for measuring
agreement among ranked lists.  The general idea is to define agreement
based on the sequence of ranks from a subset of the first $d$ items in
each list. As agreement metric we adapt the limits of agreement known
from agreement between quantitative variables
\citep{alt:bland:1983,Carstensen2010} but any measure for agreement
could essentially be used. Our proposed approach allows us to compare
multiple lists simultaneously, it provides a dynamic measure of
agreement as a function of the depth in the lists, it places higher
weight on items at the top of the list, it accommodates partially
observed lists of varying lengths, and has a natural interpretation
that directly relates to the ranks.  Graphical illustration of
sequential rank agreement allows us to infer any changepoints, i.e., a
list depth where a substantial change in the agreement of the lists
occur but we also provide asymptotical and randomization-based
graphical tools to compare the observed rank agreement to the expected
agreement found in non-informative data.  In this sense our approach
is a combination and generalization of some of the ideas of
\added{\citet{iriz:etal:2005}}, \citet{Carterette2009}, and
\citet{Boulesteix2009}. \citet{Carterette2009} compares two rankings
based on the distance between them as measured by a multivariate
Gaussian distribution and the latter presents an overview of
approaches for aggregation of ranked lists including bootstrap and
leave-one-out jackknife approaches. \added{\citet{iriz:etal:2005}
  proposes a plot which is a special case of our setting where the
  agreement metric is the overlap proportion}. We show the asymptotic
distribution of the endpoint of agreement and discuss how to infer the
distribution in small-sample situations using computational
methods. This enables us to make inferences about the endpoint of
agreement even for situations where there is no actual changepoint
present but just a gradual decline in agreement.

@@

Other recent approaches consider the intersection of lists as the
basis for a similarity measure \citep{iriz:etal:2005}. However, simple
intersection also places equal weights on all depths of the list and
therefore \citet{Fagin2003}, and \citet{Webber2010} proposed weighted
intersections which put more emphasis on the top of the lists.
% One proposal is denoted the average overlap (AO) and used for
% comparison later on. 
Specifically, \citet{Webber2010} define their rank-biased overlap
(RBO) by weighting with a converging series to ensure that the top is
weighted higher than the potentially non-informative bottom of the
lists.  It is possible to use the existing methods to calculate
agreement of lists until a given depth, i.e., limited to the $d$ items
of each list. However, the interpretation may not be straightforward,
especially in the case of more than two lists, and they may not
accommodate partial rankings.
